Positional surprisal (1-3grams) in CHILDES corpora

### 2019-09-17 surprisal/childes_surprisal.Rmd
laplace smoothing in childes surprisal

```{r hard-coded language unigrams}
child_utterances <- palasis_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_unigrams <- child_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

child <- child_utterances %>%
  left_join(child_unigrams) %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

child_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))

@@

```{r bigrams}
child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x/n.y) %>%
  select(-n.y, -p)

bigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(child_bigrams, by = c("lag_word", "word")) %>%
  left_join(child_unigrams, by = "word")

child <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

>>>>

```{r hard-coded language unigrams}
child_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_unigrams <- child_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(n = n + 1) %>%
  mutate(p = n / sum(n))

child <- child_utterances %>%
  left_join(child_unigrams, by = "word") %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  filter(length %in% c(5, 7, 9)) %>%
  group_by(length, word_order) %>%
  nest()

child %>%
  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
  select(-data) %>%
  unnest(cols = c(surprisal))

  ggplot(aes(x = word_order, y = surprisal, color = as.factor(length))) + 
  geom_point() +
  geom_line()

child_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 100)))

@@

```{r bigrams}
vocab_size <- child_utterances %>%
  ungroup() %>%
  distinct( word) %>% 
  nrow()

child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = (n.x + 0)  / (n.y + 0 * vocab_size)) %>%
  select(-n.y, -p)

bigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(child_bigrams, by = c("lag_word", "word")) #%>%
  #left_join(child_unigrams, by = "word")

child <- bigram_surprisals_prep %>%
  filter(!is.na(lag_word)) %>%
  mutate(s = -log(cond_p)) %>%
 # mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  filter(length %in% c(5, 7, 9)) %>%
  group_by(length, word_order) %>%
  nest()

child %>%
  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
  select(-data) %>%
  unnest(cols = surprisal) %>%
  ggplot(aes(x = word_order, y = empirical_stat, color = as.factor(length))) + 
  geom_point() +
  geom_line()
